Haploid Selection, Sex Ratio Bias, and Transitions Between Sex-Determination Systems

### 2017-10-25 LaTeX/Turnover.tex
minor changes, just reading

We find that tightly sex-linked genetic variation can favour the spread of new sex-determination systems in which the heterogametic sex changes (XY to ZW or ZW to XY) and the new sex-determining region is less closely linked (or unlinked) to the sex-linked locus under selection, which would not be predicted by previous theory. 

@@

The evolution of a new sex-determination system requires that a rare mutant allele at the novel sex-determining locus, $m$, increases in frequency when rare. 
The spread of a rare mutant $m$ at the \textbf{M} locus is determined by the leading eigenvalue, $\lambda$, of the system of eight equations describing the frequency of eggs and sperm carrying the $m$ allele in the next generation (equations \ref{eq:recursions}). %, evaluated at the above equilibrium.
%Thus, the system describing the change in frequency of the $m$ allele consists of eight recursion equations. 
This system simplifies substantially in a number of cases of interest. 
Dominant neo-Y (when $k=0$) or neo-W alleles (when $k=1$) are only found in male diploids (neo-Y) or female diploids (neo-W) such that their growth rate ultimately depends only on the change in frequency of $m$-bearing gametes produced by males or by females, respectively. 
Furthermore, if the $m$ allele is fully epistatically dominant over the ancestral sex-determining system, phenotypes are not affected by the genotype at the ancestral sex-determining region (\textbf{X} locus). 
Thus, the invasion of rare dominant neo-Y or neo-W alleles is determined by the largest eigenvalue that solves a quadratic characteristic polynomial,
%\textcolor{red}{Mention the possibility that the other roots yield the leading eigenvalue somewhere.}
%\textcolor{blue}{following the logic above, it seems like we say we go from 8 eqns to 4 eqns by considering only one sex, and then from 4 eqns to 2 eqns by ignoring the ancestral SD allele. why then do we have the two quadratic problem? (b/c the $m$ alleles can arise in either sex?)}
$\lambda^2+ b \lambda + c =0$ \textcolor{blue}{(see Appendix for a discussion of other roots)}.
%\begin{equation}
%\lambda^2+ b \lambda + c =0
%\label{eq:charpoly_neoY}
%\end{equation}
%
%\noindent
%where $b$ is the average of the growth rates of the two haplotypes that carry the $m$ allele ($mA$ and $ma$), $b=(\lambda_{mA}+\lambda_{ma})/2$, and $c$ also involves the fitness of $m$ alleles when they recombine onto the other \textbf{A} background in a heterozygote, $c=\lambda_{mA}\lambda_{ma}+\chi_{mA}\chi_{ma}$ (see table \ref{tab:haplotype_growth}).
%Here, $b= - (\lambda_{mA} + \lambda_{ma})$ and $c = \lambda_{mA} \lambda_{ma} -\chi_{mA} \chi_{ma}$, where $\lambda_{mi}$ is the (multiplicative) growth rate of mutant haplotypes on background $i\in\{A,a\}$, accounting for loss due to recombination, and $\chi_{mi}$ is the rate of addition of mutant haplotypes onto background $i\in\{A,a\}$ due to recombination (see table \ref{tab:haplotype_growth}).
Here, $b= - (\lambda_{mA} + \lambda_{ma})+(\chi_{mA} + \chi_{ma})$ and $c = (\lambda_{mA}-\chi_{mA}) (\lambda_{ma}-\chi_{ma}) -\chi_{mA} \chi_{ma}$, where $\lambda_{mi}$ is the multiplicative growth rate of mutant haplotypes on background $i\in\{A,a\}$, without accounting for loss due to recombination, and $\chi_{mi}$ is the rate at which mutant haplotypes on background $i\in\{A,a\}$ recombine onto the other \textbf{A} locus background in heterozygotes (see Table \ref{tab:haplotype_growth}).
The $\lambda_{mi}$ and $\chi_{mi}$, and thus the spread of the mutant $m$ allele, depend on the frequency of alleles at the \textbf{A} and \textbf{X} loci in the ancestral population. 
In the ancestral population, it is convenient to follow the frequency of the $A$ allele among female gametes (eggs), $p^\female_X$, and among X-bearing, $p^\male_X$, and among Y-bearing, $p^\male_Y$, male gametes (sperm/pollen). 
We also track the fraction of male gametes that are Y-bearing, $q$, which may deviate from $1/2$ due to meiotic drive in males. 
We will consider only equilibrium frequencies of alleles, $\hat{p}^\Hermaphrodite_i$, and Y-bearing male gametes, $\hat{q}$, to ensure the eigenvalues of the invasion analysis are valid.

@@

Assuming that linkage is not tight, \citet{vanDoorn:2010hu} showed that invasion by a neo-W occurs under the same conditions as `fixation' (where fixation indicates that the neo-W reaches its maximum frequency among eggs, which is 1/2). 
An equivalent analysis is not possible where we assume that linkage is tight. 
However, numerical simulations with tight linkage demonstrate that the neo-SDR does not necessarily fix, leading to the stable maintenance of a mixed sex-determining system, in which X,Y,Z, and W alleles all segregate, e.g., Figure \ref{fig:freqAll}B-D. 
Within a species, both feminizing and masculinizing alleles have been reported in houseflies \citep{Macdonald1978}, midges \citep{Thompson1971}, frogs \citep{Ogata:2007jm}, cichlid fish \citep{Ser:2010iq}, tilapia \citep{Lee2004}, sea bass \citep{Vandeputte2007}, and lab-strains of Zebrafish \citep{Liew2012,Wilson2014}. 
For example, in the platyfish (\textit{Xiphophorus maculatus}), X,Y, and W alleles segregate at one locus (or two closely-linked loci) near to potentially sexually-antagonistic genes for pigmentation and sexual maturity \citep{Kallman1965,Kallman1968, Volff2001, Schultheis2006}. 
Our results suggest that several forms of selection on nearby loci ($r$ and $R$ small) could maintain multiple sex determination alleles.

%Haploid selection impacts the spread of neo-W haplotypes through its direct selective effect on females and their gametes, its indirect effect on the marginal fitness of alleles in females resulting from haploid selection in male gametes, and its effect on the sex ratio.
%These impacts expand the scenarios under which neo-W haplotypes can spread.
%For instance, overdominance is no longer required for neo-W-$a$ haplotypes to spread when there is haploid selection (e.g., when $a$ is female-beneficial, $A$ is male-beneficial, and meiotic drive in males favours $a$, Figure \ref{fig:regionMaleDrive}B).

@@

Secondly, we can simplify the discussion of cases where invasion occurs despite looser sex-linkage, $R>r$, by focussing on the special case where $R=1/2$ and $r<1/2$ (e.g., the selected locus is on the ancestral sex chromosome and the novel sex-determining locus arises on an autosome). 
In table \ref{tab:specialcases} we give the conditions where invasion occurs when we further assume that haploid selection only occurs in one sex (e.g., during male meiosis only) and dominance coefficients are equal in the two sexes, $h^\female=h^\male$. 
When there is no gametic competition and meiotic drive is in one sex only, an unlinked neo-W can invade as long as the same allele is favoured during diploid selection in males and females ($s^\female s^\male>0$, see Figure \ref{fig:Combination_Centimorgans}A and Figure \ref{fig:SexRatioBad}B). %, which is 50\% of the parameter space. 
When there is no meiotic drive and gametic competition occurs in one sex only, an unlinked neo-W can invade as long as the same allele is favoured in male and female diploid selection and there are sex differences in selection of one type (e.g., $s^\female(s^\male-s^\female)>0$, see Figure \ref{fig:Combination_Centimorgans}B). %, which is 25\% of the parameter space. 
These special cases indicate that neo-W invasion occurs for a relatively large fraction of the parameter space, even if the neo-W uncouples the sex-determining locus from a locus under selection.

@@

%Under Fisherian sex-ratio selection, autosomal modifiers favour equal investment in male and female offspring, i.e., a 1:1 sex ratio \citep{Fisher:1930wy,Charnov:1982wg,West:2009we}. 
%A novel environmental sex-determiner that causes half of its carriers to become female and half to become male ($k=1/2$) will be in males half of the time and in females half of the time (like an autosome).
%In addition, these novel sex-determination alleles equalize the sex ratio and therefore one might expect them to be favoured by Fisherian sex-ratio selection when the resident sex ratio is biased.
%However, assuming weak selection, we find that the growth rate of a rare, dominant offspring-controlled neo-ESD allele that produces males or females with equal probability ($k=1/2$) is
Of particular interest are ESD mutations that cause half of their carriers to develop as females and half as males ($k=1/2$, creating equal sex ratios), the spread of which is given by
\begin{equation}
\lambda_{ESD',XY} =1+ \frac{1}{2}\frac{(\lambda_{Y',XY\rvert R=1/2}-1) + (\lambda_{W',XY\rvert R=1/2}-1)}{2} + O\left(\epsilon^3\right),
%V_A \frac{r}{1-2r}  \frac{s_{f} \left( s^\male - 3 s^\female\right) }{8 \left( s_{f} + s_{m} \right)^2} t_{m}^2.
\label{eq:lambda_ESD}
\end{equation}
\noindent
where $\lambda_{Y',XY\rvert R=1/2}$ and $\lambda_{W',XY\rvert R=1/2}$ represent $\lambda_{Y',XY}$ and $\lambda_{W',XY}$ when evaluated at $R=1/2$ (equations \ref{eq:lambda_neoY} and \ref{eq:lambda_neoW}).
That is, recombination between the selected locus and the novel sex-determining locus, $R$, doesn't enter into the $k=1/2$ results.
This is because sex is essentially randomized each generation, %as if the novel sex-determiner were on an unlinked site that half the time acts as a neo-$W$ and half the time as a neo-$Y$.
preventing associations from building up between allele $A$ and sex. 
Equation \eqref{eq:lambda_ESD} shows that the neo-ESD gets half of the fitness of a feminizing mutation (neo-$W$) and half of the fitness of a masculinizing mutation (neo-$Y$), but only has an effect one half of the time (the other half of the time it produces the same sex as the ancestral system would have, to leading order). 
As discussed above, $\lambda_{Y',XY\rvert R=1/2}$ is necessarily less than one, but $\lambda_{W',XY\rvert R=1/2}$ can be greater than one if there is haploid selection.
That is, when there is haploid selection, ESD mutations can invade an ancestrally-XY system because they generate females that are either rare or have high fitness, in the same manner as a neo-W.

@@

%COMPARATIVE PATTERNS 
We have shown that the spread of new sex determination systems can be driven by loci experiencing haploid selection. 
Because haploid selection can cause transitions that increase or decrease sex-linkage, haploid selection may lead to less stability, and greater potential for cycling, in sex-determination systems (e.g., the final state of the red line in Figure \ref{fig:SexRatioBad}A is the starting state in Figure \ref{fig:SexRatioBad}B). 
In particular, if haploid selection is strong but selective differences between male and female diploids are weak, we find that heterogametic transitions (XY to ZW or vice versa) are favoured more strongly than homogametic transitions (e.g., with $|D^\male - D^\female| << |\alpha_\Delta^\male - \alpha_\Delta^\female + t^\male - t^\female|$ we have $\lambda_{W',XY} > \lambda_{Y',XY}$; equations \ref{eq:lambda_neoW} and \ref{eq:freq_diffs}). 
Turnovers driven by haploid selection may help to explain the relative rarity of heteromorphic sex chromosomes in plants, which are thought to experience more selection during their multicellular haploid stage.
For example, among relatively few dioecious clades in which multiple species have well characterized sex chromosomes \citep{Ming:2011iy}, heterogametic transitions have been inferred in \textit{Silene} subsection \textit{Otites} \citep{Slancarova:2013dq} and in \textit{Salicaceae} \citep{Pucholt2015,Pucholt2017}.
%Potentially, successive heterogametic transitions between master regulators of sex-determination could be inferred from careful examination of the molecular pathways by which sex is determined. 
%Our predictions could also be examined using a suitable proxy for haploid selection, for example, \cite{Lenormand:2005vb} use the outcrossing rate in plants as a proxy for the strength of pollen competition. 
Furthermore, assuming that transitions from dioecy to hermaphroditism (equal parental investment in male and female gametes) are favoured in a similar manner to the ESD examined here (equal probability of zygotes developing as males or females), our results suggest that competition during the haploid stage could drive transitions between dioecy and hermaphroditism, which are frequent in plants \citep{Kafer2017, Goldberg2017}.
%In animals, one might expect gametic competition to be stronger in species where sperm is required to live for a long time after spermatogenesis because transcripts shared during spermatogenesis may become depleted, revealing the haploid phenotype of the sperm \citep{Immler:2014im}.
%We would also expect that heterogametic transitions in sex determination would be more common in clades where there is meiotic drive.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Both neo-W haplotypes can spread, in the absence of haploid selection, when there is overdominance
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!h]
\centering
\centerline{\includegraphics[width=\linewidth]{PositionPlot_Overdominance_Mike}}
\caption{
Neo-W alleles can spread when loci under diploid selection are tightly linked to the ancestral sex determining locus ($r\approx0$). 
In panels A and B, the $a$ allele is favoured in females ($w_{aa}^\female=1.05$, $w_{Aa}^\Hermaphrodite=1$, $w_{AA}^\female=0.85$) and selection in males is overdominant ($w_{aa}^\male=w_{AA}^\male=0.75$).
In panels C and D, selection in males and females is overdominant ($w_{aa}^\female=w_{AA}^\female=0.6$, $w_{aa}^\male=0.5$, $w_{AA}^\male=0.7$, $w_{Aa}^\Hermaphrodite=1$). 
These parameters are marked by a dagger in Figure \ref{fig:regionplots}, which shows that neo-W invasion is expected for any $R$ when the $a$ allele is nearly fixed on the Y (black lines). 
Equilibria where the $A$ allele is more common among Y-bearing male gametes can also be stable for these parameters (blue lines). 
The weak selection approximation holds when all recombination rates are large relative to selection (around 0 in panels A and C), in which case neo-W alleles should spread if they are more tightly linked to the selected locus (positive invasion fitness in the grey region). 
However, when linkage is tight (panels C and D and when the selected locus is near the SDR), this prediction breaks down. 
Here, there is no haploid selection $t^\Hermaphrodite = \alpha^\Hermaphrodite_\Delta = 0$.
}
\label{fig:positionOverdominance}
\end{figure}

@@

\begin{figure}[!h]
\centering
\centerline{\includegraphics[width=1.5\linewidth]{PositionPlot_SexAntagTighter_MaleDrive_Mike}}
\caption{
When there is sexually-antagonistic selection and haploid selection, a neo-W may invade for any $R$.
\textcolor{red}{
Check that we mention Sally's result that invasion cannot occur with sexually-antagonistic selection and $R=1/2$ (currently only in legend for figure \ref{fig:SexAntagTighter}). 
%Add asterisk to Figure \ref{fig:regionMaleDrive}A to show parameters used in this plot. 
}
\textcolor{blue}{This is now discussed in a round-about way at the beginning of the tight linkage invasion section in the main text. }
Parameters given by asterisk in Figure \ref{fig:regionMaleDrive}: $w_{AA}^\female=1.05$, $w_{Aa}^\Hermaphrodite=1$, $w_{aa}^\female=0.85$, $w_{AA}^\male=0.85$, $w_{aa}^\male=1.05$, $\alpha_{\Delta}^\male=-0.08$.
}
\label{fig:SexAntagTighterMaleDrive}
\end{figure}

>>>>

We find that tightly sex-linked genetic variation can favour the spread of new sex-determination systems in which the heterogametic sex changes (XY to ZW or ZW to XY) and the new sex-determining region is less closely linked (or unlinked) to the sex-linked locus under selection, which is not predicted by previous theory. 

@@

The evolution of a new sex-determination system requires that a rare mutant allele at the novel sex-determining locus, $m$, increases in frequency when rare. 
The spread of a rare mutant $m$ at the \textbf{M} locus is determined by the leading eigenvalue, $\lambda$, of the system of eight equations describing the frequency of eggs and sperm carrying the $m$ allele in the next generation (equations \ref{eq:recursions}). %, evaluated at the above equilibrium.
%Thus, the system describing the change in frequency of the $m$ allele consists of eight recursion equations. 
This system simplifies substantially in a number of cases of interest. 
Dominant neo-Y (when $k=0$) or neo-W alleles (when $k=1$) are only found in male diploids (neo-Y) or female diploids (neo-W) such that their growth rate ultimately depends only on the change in frequency of $m$-bearing gametes produced by males or by females, respectively. 
Furthermore, if the $m$ allele is fully epistatically dominant over the ancestral sex-determining system, phenotypes are not affected by the genotype at the ancestral sex-determining region (\textbf{X} locus). 
Thus, the invasion of rare dominant neo-Y or neo-W alleles is determined by the largest eigenvalue that solves a quadratic characteristic polynomial,
%\textcolor{red}{Mention the possibility that the other roots yield the leading eigenvalue somewhere.}
%\textcolor{blue}{following the logic above, it seems like we say we go from 8 eqns to 4 eqns by considering only one sex, and then from 4 eqns to 2 eqns by ignoring the ancestral SD allele. why then do we have the two quadratic problem? (b/c the $m$ alleles can arise in either sex?)}
$\lambda^2+ b \lambda + c =0$ \textcolor{blue}{(see Appendix for a discussion of other roots - or Sally's proof!)}.
%\begin{equation}
%\lambda^2+ b \lambda + c =0
%\label{eq:charpoly_neoY}
%\end{equation}
%
%\noindent
%where $b$ is the average of the growth rates of the two haplotypes that carry the $m$ allele ($mA$ and $ma$), $b=(\lambda_{mA}+\lambda_{ma})/2$, and $c$ also involves the fitness of $m$ alleles when they recombine onto the other \textbf{A} background in a heterozygote, $c=\lambda_{mA}\lambda_{ma}+\chi_{mA}\chi_{ma}$ (see table \ref{tab:haplotype_growth}).
%Here, $b= - (\lambda_{mA} + \lambda_{ma})$ and $c = \lambda_{mA} \lambda_{ma} -\chi_{mA} \chi_{ma}$, where $\lambda_{mi}$ is the (multiplicative) growth rate of mutant haplotypes on background $i\in\{A,a\}$, accounting for loss due to recombination, and $\chi_{mi}$ is the rate of addition of mutant haplotypes onto background $i\in\{A,a\}$ due to recombination (see table \ref{tab:haplotype_growth}).
Here, $b= - (\lambda_{mA} + \lambda_{ma})+(\chi_{mA} + \chi_{ma})$ and $c = (\lambda_{mA}-\chi_{mA}) (\lambda_{ma}-\chi_{ma}) -\chi_{mA} \chi_{ma}$, where $\lambda_{mi}$ is the multiplicative growth rate of mutant haplotypes on background $i\in\{A,a\}$, without accounting for loss due to recombination, and $\chi_{mi}$ is the rate at which mutant haplotypes on background $i\in\{A,a\}$ recombine onto the other \textbf{A} locus background in heterozygotes (see Table \ref{tab:haplotype_growth}).
The $\lambda_{mi}$ and $\chi_{mi}$, and thus the spread of the mutant $m$ allele, depend on the frequency of alleles at the \textbf{A} and \textbf{X} loci in the ancestral population. 
In the ancestral population, it is convenient to follow the frequency of the $A$ allele among female gametes (eggs), $p^\female_X$, and among X-bearing, $p^\male_X$, and among Y-bearing, $p^\male_Y$, male gametes (sperm/pollen). 
We also track the fraction of male gametes that are Y-bearing, $q$, which may deviate from $1/2$ due to meiotic drive in males. 
We will consider only equilibrium frequencies of alleles, $\hat{p}^\Hermaphrodite_i$, and Y-bearing male gametes, $\hat{q}$, to ensure the eigenvalues of the invasion analysis are valid.

@@

Assuming that linkage is not tight, \citet{vanDoorn:2010hu} showed that invasion by a neo-W occurs under the same conditions as `fixation' (where fixation indicates that the neo-W reaches its maximum frequency among eggs, which is 1/2). 
An equivalent analysis is not possible where we assume that linkage is tight. 
However, numerical simulations with tight linkage demonstrate that the neo-SDR does not necessarily fix, leading to the stable maintenance of a mixed sex-determining system, in which X, Y, Z, and W alleles all segregate (e.g., Figure \ref{fig:freqAll}B-D). 
Within a species, both feminizing and masculinizing alleles have been reported in houseflies \citep{Macdonald1978}, midges \citep{Thompson1971}, frogs \citep{Ogata:2007jm}, cichlid fish \citep{Ser:2010iq}, tilapia \citep{Lee2004}, sea bass \citep{Vandeputte2007}, and lab-strains of Zebrafish \citep{Liew2012,Wilson2014}. 
For example, in the platyfish (\textit{Xiphophorus maculatus}), X,Y, and W alleles segregate at one locus (or two closely-linked loci) near to potentially sexually-antagonistic genes for pigmentation and sexual maturity \citep{Kallman1965,Kallman1968, Volff2001, Schultheis2006}. 
Our results suggest that several forms of selection on nearby loci (i.e., $r$ and $R$ small) could maintain multiple sex-determining alleles.

%Haploid selection impacts the spread of neo-W haplotypes through its direct selective effect on females and their gametes, its indirect effect on the marginal fitness of alleles in females resulting from haploid selection in male gametes, and its effect on the sex ratio.
%These impacts expand the scenarios under which neo-W haplotypes can spread.
%For instance, overdominance is no longer required for neo-W-$a$ haplotypes to spread when there is haploid selection (e.g., when $a$ is female-beneficial, $A$ is male-beneficial, and meiotic drive in males favours $a$, Figure \ref{fig:regionMaleDrive}B).
\textcolor{blue}{haploid selection?}

@@

Secondly, we can simplify the discussion of cases where invasion occurs despite looser sex-linkage, $R>r$, by focusing on the special case where $R=1/2$ and $r<1/2$ (e.g., the selected locus is on the ancestral sex chromosome and the novel sex-determining locus arises on an autosome). 
In Table \ref{tab:specialcases} we give the conditions where invasion occurs when we further assume that haploid selection only occurs in one sex (e.g., during male meiosis only) and dominance coefficients are equal in the two sexes, $h^\female=h^\male$. 
When there is no gametic competition and meiotic drive is in one sex only, an unlinked neo-W can invade as long as the same allele is favoured during diploid selection in males and females ($s^\female s^\male>0$, see Figure \ref{fig:Combination_Centimorgans}A and Figure \ref{fig:SexRatioBad}B). %, which is 50\% of the parameter space. 
When there is no meiotic drive and gametic competition occurs in one sex only, an unlinked neo-W can invade as long as the same allele is favoured in male and female diploid selection and there are sex differences in selection of one type (e.g., $s^\female(s^\male-s^\female)>0$, see Figure \ref{fig:Combination_Centimorgans}B). %, which is 25\% of the parameter space. 
These special cases indicate that neo-W invasion occurs for a relatively large fraction of the parameter space, even if the neo-W uncouples the sex-determining locus from a locus under selection.

@@

%Under Fisherian sex-ratio selection, autosomal modifiers favour equal investment in male and female offspring, i.e., a 1:1 sex ratio \citep{Fisher:1930wy,Charnov:1982wg,West:2009we}. 
%A novel environmental sex-determiner that causes half of its carriers to become female and half to become male ($k=1/2$) will be in males half of the time and in females half of the time (like an autosome).
%In addition, these novel sex-determination alleles equalize the sex ratio and therefore one might expect them to be favoured by Fisherian sex-ratio selection when the resident sex ratio is biased.
%However, assuming weak selection, we find that the growth rate of a rare, dominant offspring-controlled neo-ESD allele that produces males or females with equal probability ($k=1/2$) is
Of particular interest are ESD mutations that cause half of their carriers to develop as females and half as males ($k=1/2$, creating equal sex ratios), the spread of which is given by
\begin{equation}
\lambda_{ESD',XY} =1+ \frac{1}{2}\frac{(\lambda_{Y',XY\rvert R=1/2}-1) + (\lambda_{W',XY\rvert R=1/2}-1)}{2} + O\left(\epsilon^3\right),
%V_A \frac{r}{1-2r}  \frac{s_{f} \left( s^\male - 3 s^\female\right) }{8 \left( s_{f} + s_{m} \right)^2} t_{m}^2.
\label{eq:lambda_ESD}
\end{equation}
\noindent
where $\lambda_{Y',XY\rvert R=1/2}$ and $\lambda_{W',XY\rvert R=1/2}$ represent $\lambda_{Y',XY}$ and $\lambda_{W',XY}$ when evaluated at $R=1/2$ (Equations \ref{eq:lambda_neoY} and \ref{eq:lambda_neoW}).
That is, recombination between the selected locus and the novel sex-determining locus, $R$, doesn't enter into the $k=1/2$ results.
This is because sex is essentially randomized each generation, %as if the novel sex-determiner were on an unlinked site that half the time acts as a neo-$W$ and half the time as a neo-$Y$.
preventing associations from building up between allele $A$ and sex. 
Equation \eqref{eq:lambda_ESD} shows that the neo-ESD gets half of the fitness of a feminizing mutation (neo-$W$) and half of the fitness of a masculinizing mutation (neo-$Y$), but only has an effect one half of the time (the other half of the time it produces the same sex as the ancestral system would have, to leading order). 
As discussed above, $\lambda_{Y',XY\rvert R=1/2}$ is necessarily less than one, but $\lambda_{W',XY\rvert R=1/2}$ can be greater than one if there is haploid selection.
That is, when there is haploid selection, ESD mutations can invade an ancestrally-XY system because they generate females that are either rare or have high fitness, in the same manner as a neo-W.

@@

%COMPARATIVE PATTERNS 
We have shown that the spread of new sex determination systems can be driven by loci experiencing haploid selection. 
Because haploid selection can cause transitions that increase or decrease sex-linkage, haploid selection may lead to less stability, and greater potential for cycling, in sex-determination systems (e.g., the final state of the red line in Figure \ref{fig:SexRatioBad}A is the starting state in Figure \ref{fig:SexRatioBad}B). 
In particular, if haploid selection is strong but selective differences between male and female diploids are weak, we find that heterogametic transitions (XY to ZW or vice versa) are favoured more strongly than homogametic transitions (e.g., with $|D^\male - D^\female| << |\alpha_\Delta^\male - \alpha_\Delta^\female + t^\male - t^\female|$ we have $\lambda_{W',XY} > \lambda_{Y',XY}$; Equations \ref{eq:lambda_neoW} and \ref{eq:freq_diffs}). 
Turnovers driven by haploid selection may help to explain the relative rarity of heteromorphic sex chromosomes in plants, which are thought to experience more selection during their multicellular haploid stage.
For example, among relatively few dioecious clades in which multiple species have well characterized sex chromosomes \citep{Ming:2011iy}, heterogametic transitions have been inferred in \textit{Silene} subsection \textit{Otites} \citep{Slancarova:2013dq} and in \textit{Salicaceae} \citep{Pucholt2015,Pucholt2017}.
%Potentially, successive heterogametic transitions between master regulators of sex-determination could be inferred from careful examination of the molecular pathways by which sex is determined. 
%Our predictions could also be examined using a suitable proxy for haploid selection, for example, \cite{Lenormand:2005vb} use the outcrossing rate in plants as a proxy for the strength of pollen competition. 
Furthermore, assuming that transitions from dioecy to hermaphroditism (equal parental investment in male and female gametes) are favoured in a similar manner to the ESD examined here (equal probability of zygotes developing as males or females), our results suggest that competition during the haploid stage could drive transitions between dioecy and hermaphroditism, which are frequent in plants \citep{Kafer2017, Goldberg2017}.
%In animals, one might expect gametic competition to be stronger in species where sperm is required to live for a long time after spermatogenesis because transcripts shared during spermatogenesis may become depleted, revealing the haploid phenotype of the sperm \citep{Immler:2014im}.
%We would also expect that heterogametic transitions in sex determination would be more common in clades where there is meiotic drive.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Both neo-W haplotypes can spread, in the absence of haploid selection, when there is overdominance
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!h]
\centering
\centerline{\includegraphics[width=\linewidth]{PositionPlot_Overdominance_Mike}}
\caption{
Neo-W alleles can spread when loci under diploid selection are tightly linked to the ancestral sex determining locus ($r\approx0$). 
In panels A and B, the $a$ allele is favoured in females ($w_{aa}^\female=1.05$, $w_{Aa}^\Hermaphrodite=1$, $w_{AA}^\female=0.85$) and selection in males is overdominant ($w_{aa}^\male=w_{AA}^\male=0.75$).
In panels C and D, selection in males and females is overdominant ($w_{aa}^\female=w_{AA}^\female=0.6$, $w_{aa}^\male=0.5$, $w_{AA}^\male=0.7$, $w_{Aa}^\Hermaphrodite=1$).
There is no haploid selection $t^\Hermaphrodite = \alpha^\Hermaphrodite_\Delta = 0$.
These parameters are marked by daggers in Figure \ref{fig:regionplots}B and C, which show that neo-W invasion is expected for any $R$ ($\lambda_{WA},\lambda_{Wa}>1$) when the $a$ allele is nearly fixed on the Y (black lines in this figure). 
Equilibria where the $A$ allele is more common among Y-bearing male gametes can also be stable and allow neo-W invasion for these parameters (blue lines). 
The weak selection approximation holds when all recombination rates are large relative to selection (around 0 in panels A and C), in which case, in the absence of haploid selection, neo-W alleles should spread only if they are more tightly linked to the selected locus (positive invasion fitness if and only if the selected locus is in the grey region). 
However, when linkage is tight (panels C and D and when the selected locus is near the SDR in all panels), this prediction breaks down. 
}
\label{fig:positionOverdominance}
\end{figure}

@@

\begin{figure}[!h]
\centering
\centerline{\includegraphics[width=1.5\linewidth]{PositionPlot_SexAntagTighter_MaleDrive_Mike}}
\caption{
When there is sexually-antagonistic selection and haploid selection, a neo-W may invade for any $R$.
\textcolor{red}{
Check that we mention Sally's result that invasion cannot occur with sexually-antagonistic selection and $R=1/2$ (currently only in legend for figure \ref{fig:SexAntagTighter}). 
%Add asterisk to Figure \ref{fig:regionMaleDrive}A to show parameters used in this plot. 
}
\textcolor{blue}{This is now discussed in a round-about way at the beginning of the tight linkage invasion section in the main text. }
Parameters given by asterisk in Figure \ref{fig:regionMaleDrive}A: $w_{AA}^\female=1.05$, $w_{Aa}^\Hermaphrodite=1$, $w_{aa}^\female=0.85$, $w_{AA}^\male=0.85$, $w_{aa}^\male=1.05$, $\alpha_{\Delta}^\male=-0.08$.
}
\label{fig:SexAntagTighterMaleDrive}
\end{figure}
